Automatically generated NE tagged corpora for English and Hungarian

### 2012-04-26 paper/engNERwiki.tex

Although Hungarian is reckoned to be a less resourced language, and it is not supperted in NLTK, several high quality language processing tools have been developed for Hungarian in recent years. For tokenization and sentence segmentation we used an in-house statistical tool tailored for Hungarian. It has been trained on the largest manually annotated Hungarian corpus \cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions. Lemmatization was performed by \texttt{hunmorph} \cite{Tron:05} and \texttt{hundisambig}, the latter being an in-house disambiguator to select the right analysis based on the word context.

@@

For well-resourced languages, DBpedia has internationalized chapters; data dumps based on Wikipedia. Hungarian does not have a separate chapter. Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia. Two consequences follow.

@@

The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as NEs. As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized. The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs according to the Szeged NER guideline: names of languages, nationalities, religions, political ideologies; adjectives derived from a NE; names of months, days, holidays; names of special events and wars.

>>>>

Although Hungarian is reckoned to be a less resourced language, and it is not supperted in NLTK, several high quality language processing tools have been developed for Hungarian in recent years. For tokenization and sentence segmentation we used an in-house statistical tool tailored for Hungarian. It has been trained on the largest manually annotated Hungarian corpus \cite{Csendes:04}, and it handles the peculiarities of Hungarian ortography, such as the periods placed after numbers in date expressions. Lemmatization was performed by hunmorph \cite{Tron:05} and hundisambig, the latter being an in-house disambiguator to select the right analysis based on the word context.

@@

For well-resourced languages, DBpedia has internationalized chapters, 
%data dumps based on Wikipedia. 
but not for Hungarian. Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia. Two consequences follow.

@@

The Szeged NER guideline relies heavily on the rules of capitalization to decide which words should be marked as NEs. As mentioned earlier, in Hungarian, apart from the first word in the sentence, only proper nouns are capitalized. The following concepts are not proper nouns in Hungarian, and thus are not considered as NEs: names of languages, nationalities, religions, political ideologies; adjectives derived from a NE; names of months, days, holidays; names of special events and wars.
